---
title: "STAT/MATH 495: Problem Set 03"
author: "WRITE YOUR NAME HERE"
date: "2017-09-26"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    collapsed: false
    smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=4.5)

# Load packages
library(tidyverse)
data1 <- read_csv("data/data1.csv")
data2 <- read_csv("data/data2.csv")
```


# Question

For both `data1` and `data2` tibbles (a tibble is a data frame with some
[metadata](https://blog.rstudio.com/2016/03/24/tibble-1-0-0#tibbles-vs-data-frames) attached):

* Find the splines model with the best out-of-sample predictive ability.
* Create a visualizaztion arguing why you chose this particular model.
* Create a visualizaztion of this model plotted over the given $(x_i, y_i)$ points for $i=1,\ldots,n=3000$.
* Give your estimate $\widehat{\sigma}$ of $\sigma$ where the noise component $\epsilon_i$ is distributed with mean 0 and standard deviation $\sigma$.


# Data 1

```{r, echo=TRUE, warning=FALSE, message=FALSE}
```

# Data 2

```{r, echo=TRUE, warning=FALSE, message=FALSE}
```

